---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.11.3
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Advent of Code (2015)

```{r}
txt <- scan('puzzle.aoc', what='character', sep='\n', fileEncoding='utf8')
vec <- strsplit(txt, '')[[1]]
```

## Day 1: Not Quite Lisp
Santa was hoping for a white Christmas, but his weather machine's "snow" function is powered by stars, and he's fresh out! To save Christmas, he needs you to collect **fifty stars** by December 25th.

Collect stars by helping Santa solve puzzles. Two puzzles will be made available on each day in the Advent calendar; the second puzzle is unlocked when you complete the first. Each puzzle grants **one star**. Good luck!

Here's an easy puzzle to warm you up.

Santa is trying to deliver presents in a large apartment building, but he can't find the right floor - the directions he got are a little confusing. He starts on the ground floor (floor 0) and then follows the instructions one character at a time.

An opening parenthesis, (, means he should go up one floor, and a closing parenthesis, ), means he should go down one floor.

The apartment building is very tall, and the basement is very deep; he will never find the top or bottom floors.

For example:

- `(())` and `()()` both result in floor `0`.
- `(((` and `(()(()(` both result in floor `3`.
- `))(((((` also results in floor `3`.
- `())` and `))(` both result in floor `-1` (the first basement level).
- `)))` and `)())())` both result in floor `-3`.


### Part 1: ***To what floor do the instructions take Santa?***

```{r}
indices_left <- which(vec=='(')
indices_right <- which(vec==')')

# replace '(',')' with respective number
vec2 <- replace(vec, indices_left, 1)
vec2 <- replace(vec2, indices_right, -1)
vec2 <- strtoi(vec2)
```

It is this floor!

```{r}
sum(vec2)
```

### Part 2
Now, given the same instructions, find the **position** of the first character that causes him to enter the basement (floor `-1`). The first character in the instructions has position `1`, the second character has position `2`, and so on.

For example:

- `)` causes him to enter the basement at character position `1`.
- `()())` causes him to enter the basement at character position `5`.

***What is the position of the character that causes Santa to first enter the basement?***

```{r}
out <- 0; index <- 0; done <- FALSE
for (x in vec2) {
  if (out == -1 && i > 1) {
    done <- TRUE
  }
  if (!done) {
    out <- out + x
    index <- index + 1
  }
}
```

Its this Instruction!

```{r}
index
```
